Bird beak's elaboration and innovation

### 2022-10-25 Analysis/03.1-Elaboration_innovation_analysis[statistics].Rmd
Update 03.1-Elaboration_innovation_analysis[statistics].Rmd

```{r, eval = TRUE, echo = FALSE, fig.height = 20, fig.width = 8}
#TODO: make the groups nested in the fable
pdf("../Manuscript/Figures/orthogonality_results.pdf", width = 9, height = 9)
source("../Functions/ploting.utilities.R")
## Plot the orthogonality into context
orthogonality <- all_results[, c(1:2, 9:14, 3, 5)]

@@

## Sort each level by orthogonality score
ortho.sort <- function(data) {
    return(data[order(data$orthogonality, decreasing = TRUE), , drop = FALSE])
}
## Plotting table (with the sorted orthogonality)
plot_table <- rbind(ortho.sort(super_orders),
                    ortho.sort(orders))
                    # ortho.sort(sub_orders),
                    # ortho.sort(families))
colours <- c(rep(gg.color.hue(4)[1], nrow(super_orders)),
             rep(gg.color.hue(4)[2], nrow(orders)))
             # rep(gg.color.hue(4)[3], nrow(sub_orders)),
             # rep(gg.color.hue(4)[4], nrow(families)))

## Adding starts here and there
get.signif <- function(prob) {
    if(prob >= 0.99) {
        return("***")
    } else {
        if(prob >= 0.95) {
            return("**")
        } else {
            if(prob >= 0.90) {
                return("*")
            } else {
                if(prob >= 0.80) {
                    return(".")
                } else {
                    return("")
                }
            }
        }
    }
}
add.stars <- function(data, counter, shift = 0, total, ...) {
    text(data[3] + 0.03, (total+1)-counter + shift + 0.2, labels = get.signif(data[6]), cex = 1, ...)   
}
add.line <- function(data, counter, shift = 0, total, pch = 19, lty = 1, ...) {
    lines(matrix(c(data[4:5], rep((total+1)-counter + shift, 2)), 2, 2), lty = lty, ...)
    points(data[3], (total+1)-counter + shift, pch = pch,  ...)
}

## Add the number of species and sd to the group names
convert.spaces <- function(x, max) {
    # shifts <- max-nchar(as.character(x))
    # if(shifts > 0) {
    #     return(paste(c(x, rep(" ", shifts)), collapse = ""))
    # } else {
    #     return(x)
    # }
    return(x)
}
group_names <- paste0(plot_table$Group, " (n = ", sapply(plot_table$n, convert.spaces, max = 4), " ; sd = ", sapply(round(plot_table$`ellipse sd`, 2), convert.spaces, max = 5), ")")

plot_order <- 1:length(group_names)

## Plot frame
par(mar = c(5, 17, 4, 2))
plot(NULL, yaxt = "n", xaxt = "n", ylab = "", xlab = "orthogonality", ylim = c(1,nrow(plot_table)), xlim = c(0,1))
abline(v = 0.5, lwd = 0.5, col = "grey")
abline(v = 0.75, lty = 2, col = "grey", lwd = 0.5)
axis(2, at = nrow(plot_table):1, labels = group_names[plot_order], las = 2)
axis(1, at = c(0, 0.5, 1), labels = c(0, 0.5, 1))
## Adding the results
for(i in 1:nrow(plot_table)) {
    add.line (plot_table[plot_order[i], ], counter = i, col = colours[plot_order[i]], total = nrow(plot_table))
    add.stars(plot_table[plot_order[i], ], counter = i, col = colours[plot_order[i]], total = nrow(plot_table))
}

## Adding the n2 results
orders_n2_plot <- orders_n2[match(orders[order(orders$orthogonality, decreasing = TRUE), ]$Group, orders_n2$Group), ]
# families_n2_plot <- families_n2[match(families[order(families$orthogonality, decreasing = TRUE), ]$Group, families_n2$Group), ]
for(i in 1:nrow(orders_n2_plot)) {
    add.line(orders_n2_plot[i, ], shift = -1/3, counter = i+nrow(super_orders), col = adjustcolor(colours[i+nrow(super_orders)], alpha.f = 2/3), total = nrow(plot_table), lty = 2, pch = 21)
    add.stars(orders_n2_plot[i, ], shift = -1/3, counter = i+nrow(super_orders), col = adjustcolor(colours[i+nrow(super_orders)], alpha.f = 2/3), total = nrow(plot_table))
}
# for(i in 1:nrow(families_n2)) {
#     add.line(families_n2_plot[i, ], shift = -1/3, counter = i+(nrow(plot_table)-nrow(families)), col = adjustcolor(colours[i+(nrow(plot_table)-nrow(families))], alpha.f = 0.5), total = nrow(plot_table))
#     add.line(families_n2_plot[i, ], shift = -1/3, counter = i+(nrow(plot_table)-nrow(families)), col = adjustcolor(colours[i+(nrow(plot_table)-nrow(families))], alpha.f = 0.5), total = nrow(plot_table))
# }

legend("topleft", legend = c("Super orders (vs. phylogeny)", "Orders (vs. super orders)", "Orders (vs. phylogeny)"), col = c(gg.color.hue(4)[1], gg.color.hue(4)[2], adjustcolor(gg.color.hue(4)[2], alpha.f = 2/3)), pch = c(19, 19, 21), lty = c(1,1,2), cex = 0.8)

# legend("topleft", legend = c("Super orders (vs. phylogeny)", "Orders (vs. super orders)", "Orders (vs. phylogeny)", "Suborders (vs. passeriformes)", "Families (vs. suborders)", "Families (vs. passeriformes)"), col = c(gg.color.hue(4)[1], gg.color.hue(4)[2], adjustcolor(gg.color.hue(4)[2], alpha.f = 1/3), gg.color.hue(4)[3], gg.color.hue(4)[4], adjustcolor(gg.color.hue(4)[4], alpha.f = 1/3)), pch = 19, lty = 1, cex = 0.8)

dev.off()
```

@@

```{r, eval = TRUE, echo = FALSE, fig.height = 20, fig.width = 8}
## Same but for the passeriformes
pdf("../Manuscript/Figures/orthogonality_results_passeriformes.pdf", width = 9, height = 9)
source("../Functions/ploting.utilities.R")
## Plotting table (with the sorted orthogonality)
plot_table <- rbind(ortho.sort(sub_orders),
                    ortho.sort(families))
colours <- c(rep(gg.color.hue(4)[3], nrow(sub_orders)),
             rep(gg.color.hue(4)[4], nrow(families)))
group_names <- paste0(plot_table$Group, " (n = ", sapply(plot_table$n, convert.spaces, max = 4), " ; sd = ", sapply(round(plot_table$`ellipse sd`, 2), convert.spaces, max = 5), ")")
plot_order <- 1:length(group_names)

## Plot frame
par(mar = c(5, 17, 4, 2))
plot(NULL, yaxt = "n", xaxt = "n", ylab = "", xlab = "orthogonality", ylim = c(1,nrow(plot_table)), xlim = c(0,1))
abline(v = 0.5, lwd = 0.5, col = "grey")
abline(v = 0.75, lty = 2, col = "grey", lwd = 0.5)
axis(2, at = nrow(plot_table):1, labels = group_names[plot_order], las = 2)
axis(1, at = c(0, 0.5, 1), labels = c(0, 0.5, 1))
## Adding the results
for(i in 1:nrow(plot_table)) {
    add.line (plot_table[plot_order[i], ], counter = i, col = colours[plot_order[i]], total = nrow(plot_table))
    add.stars(plot_table[plot_order[i], ], counter = i, col = colours[plot_order[i]], total = nrow(plot_table))
}

## Adding the n2 results
families_n2_plot <- families_n2[match(families[order(families$orthogonality, decreasing = TRUE), ]$Group, families_n2$Group), ]
for(i in 1:nrow(families_n2_plot)) {
    add.line(families_n2_plot[i, ], shift = -1/3, counter = i+nrow(sub_orders), col = adjustcolor(colours[i+nrow(sub_orders)], alpha.f = 2/3), total = nrow(plot_table), lty = 2, pch = 21)
    add.stars(families_n2_plot[i, ], shift = -1/3, counter = i+nrow(sub_orders), col = adjustcolor(colours[i+nrow(sub_orders)], alpha.f = 2/3), total = nrow(plot_table))
}

legend("topleft", legend = c("Sub-orders (vs. passeriformes)", "Families (vs. sub-orders)", "Families (vs. passeriformes)"), col = c(gg.color.hue(4)[3], gg.color.hue(4)[4], adjustcolor(gg.color.hue(4)[4], alpha.f = 2/3)), pch = c(19, 19, 21), lty = c(1,1,2), cex = 0.8)

# legend("topleft", legend = c("Super orders (vs. phylogeny)", "Orders (vs. super orders)", "Orders (vs. phylogeny)", "Suborders (vs. passeriformes)", "Families (vs. suborders)", "Families (vs. passeriformes)"), col = c(gg.color.hue(4)[1], gg.color.hue(4)[2], adjustcolor(gg.color.hue(4)[2], alpha.f = 1/3), gg.color.hue(4)[3], gg.color.hue(4)[4], adjustcolor(gg.color.hue(4)[4], alpha.f = 1/3)), pch = 19, lty = 1, cex = 0.8)

dev.off()
```

>>>>

```{r, eval = TRUE, echo = FALSE, fig.height = 20, fig.width = 8}
#TODO: make the groups nested in the fable
source("../Functions/ploting.utilities.R")
## Plot the orthogonality into context
orthogonality <- all_results[, c(1:2, 9:14, 3, 5)]

@@

## Adding orders and families that don't have a higher level covered
if(any(no_match <- is.na(match(orders_n2$Group, orders$Group)))) {
    fill_in <- orders_n2[which(no_match), ]
    fill_in[,c(3:8)] <- NA
    orders <- rbind(orders, fill_in)
}
## Adding orders and families that don't have a higher level covered
if(any(no_match <- is.na(match(families_n2$Group, families$Group)))) {
    fill_in <- families_n2[which(no_match), ]
    fill_in[,c(3:8)] <- NA
    families <- rbind(families, fill_in)
}

pdf("../Manuscript/Figures/orthogonality_results.pdf", width = 9, height = 9)
plot.orthogonality(super_orders, orders_n2, orders, col = gg.color.hue(4)[1:2])
legend("topleft", legend = c("Super orders (vs. phylogeny)", "Orders (vs. phylogeny)", "Orders (vs. super orders)"), col = c(gg.color.hue(4)[1], gg.color.hue(4)[2], adjustcolor(gg.color.hue(4)[2], alpha.f = 2/3)), pch = c(19, 19, 21), lty = c(1,1,2), cex = 0.8)
dev.off()
```

@@

```{r, eval = TRUE, echo = FALSE, fig.height = 20, fig.width = 8}
## Same but for the passeriformes
pdf("../Manuscript/Figures/orthogonality_results_passeriformes.pdf", width = 9, height = 9)
plot.orthogonality(sub_orders, families_n2, families, col = gg.color.hue(4)[3:4])
legend("topleft", legend = c("Sub-orders (vs. passeriformes)", "Families (vs. sub-orders)", "Families (vs. passeriformes)"), col = c(gg.color.hue(4)[3], gg.color.hue(4)[4], adjustcolor(gg.color.hue(4)[4], alpha.f = 2/3)), pch = c(19, 19, 21), lty = c(1,1,2), cex = 0.8)
dev.off()
```
